Hierarchical wrinkling in a confined permeable biogel

### 2015-02-12 fig_plis/article_plis.tex
Fautes d'orthographe les plus visibles Références manquantes de l'intro casein (plural or not)

\begin{abstract}
The slow acidification of milk proteins such as casein is well known to result in the formation of gels – this is indeed one of the key steps to make cheese or yogurt. Here, using the casein proteins as  building blocks and the acidifier  Glucono delta-lactone, we design model thin bio-gel layers that wrinkles in cascade. Using a combination of titration, rheology, light microscopy and confocal microscopy, we demonstrate that, during the acidification, the gels shrink and swell creating constrains that induce Darcy and Poiseuille flows and wrinkles the gel. We have develop a hydrodynamic model that predict correctly the wrinkling wavelength. Our results should be universal for proteins as it is based on pH induced charge stabilasion/destablisation and control of boundary conditons. As such it could set a benchmark to gain fundamental insights into wrinkled biological tissues, to texture food or to design surfaces for optical puposes.
\end{abstract}

@@

There are many ways and reasons for a film to wrinkle. An elastic film buckles due to excess area with respect to its boundaries and wrinkles when buckling is hindered by a substrate. The selection of the preferred wrinkling wavelength is peculiar to the situation and is related to the mismatch of elastic properties between the surface and the substrate~\cite{Gough1940, Bijlaard1946} XXX.  On the biology side wrinkling-controlled morphogenesis is ubiquitous. Ageing and the loss of elastic fibres makes our skin wrinkles~\cite{Bissett1987, Genzer2006}. Difference in growth rates between the gut tube and its dorsal anchoring is responsible for the vilification of guts~\cite{Savin2011, Ciarletta2014, Shyer2013}. Localized cell death in biofilms focuses mechanical forces and initiates 3D labyrinth pattern~\cite{Trejo2013, Asally2012}. On the physics side the last two decades have seen a bloom of methods to obtain well-controlled patterns via linear~\cite{Bowden1998, Genzer2006, Hu1998, Kim2010, Vandeparre2011, Li2013} or nonlinear behaviour~\cite{Efimenko2005, Guvendiren2010, Kim2011a, Brau2011}. Such patterns can be triggered by temperature dilation refXXX, swelling refXXX or the removal of pre-strain refXXX.

@@

Here, we introduce model permeable biogels that produce hierarchical wrinkling as a result of their interplay between their gelation dynamics and the confinement conditions. In slit geometry, we slowly acidify a dispersion of casein, one of the milk proteins.  As a result, not only does the casein form a gel as expected~\cite{Bremer1989} --this is a well-known effect used in cheese or yogurt making-- but the gel starts to produce wrinkling patterns in cascade when casein adhesion to the slit walls is turned off. Using a combination of rheology, light microscopy and confocal microscopy, we demonstrate that upon acidification the casein network spontaneously shrinks then swells at the micron level and wrinkles at the millimetre scale. We systematically investigate such swelling-induced wrinkling of a porous gel film sandwiched between two viscous layers, the whole system being confined between two glass plates. As shown in Fig.~\ref{fig:dynamics} and in Supplementary Movies 1 and 2, transverse confinement leads to saturation of the amplitude on both the top and bottom walls followed by a cascade of secondary and ternary buckling. Such cascade buckling has already been observed by~\cite{Roman1999} in a one-dimensional macroscopic system, a sheet of steel forced to buckle between two solid plates. In addition one has to note that the primary wavelength is much smaller than the size of the system, implying that a simple buckling scenario cannot explain the observations in Fig.~\ref{fig:dynamics} and that only a wrinkling instability can be invoked.

@@

\begin{figure}
	\includegraphics{acidification}%
	\caption{Acid-induced protein gel properties behave nonmonotonically with pH. (a) pH decrease over time in a 4\%w sodium caseinate solution acidified by 4\%w GDL. The horizontal line indicates isoelectric pH of casein. (b) Evolution of the storage modulus $G'$ measured in a rheometer with full adhesion to rotor and stator. (c) Evolution of pore size measured by confocal microscopy in a slit geometry with either adhesion to all walls (black line) or no adhesion to top wall, i.e. allowed syn\ae{}resis and swelling (orange dashed line).}%
	\label{fig:acidification}%
\end{figure}

To form wrinkled patterns such as the one oberved in figure 1XXX, ones needs to take advantage of the pH sensitivity of casein to control the route to gelation, to fine tune the gel properties and the boundary conditions.

the stability of casein dispersion strongly depend on the pH of the solution. Supplementary movie 3XXX shows the titration of a casein dispertion at 1\%w by 1 molar HCl. At pH=7, the casein is well dispersed then it floculate between pH= XXX to XXX and finaly at pH=XXX the solution become clear again. However as one increases the casein concentration up to 4\%w, at low pH the casein redispersion is not total, Supplementary movie 4XXX. The state diagram of casein as a function of the pH  is related to charge stabilization. At neutral pH casein are charged negatively and repel each other. As the pH decreases to the isoelectric point of the caseine (pH=4.5), caseine molecules loose their negative net charge and tend to stick to each other [ref XXX casein stability]. Beyond the isoelectric point casein acquire a positive net charge and therefore partialy re-disolve. In figure 2cXXX, we quantify this effect and measure with UV absorption spectroscopie the amount of free casein within the 4\%w dispersion as a function of pH (see Supplementary Methods). Starting at pH=6.8 all the casein molecules are well dispersed and the free casein fraction is 1. Upon decreasing the pH of the solution to the isolectric point the free casein fraction drop to almost 0, almost all the casein molecule are aggregated. Further decrease of the pH of the solution lead to a partial redisolution of the casein agregate. At pH=2 the fraction of free casein is XXX.

@@

To form a space spaning gel, floculation need to be prevented. We use of Glucono delta-lactone (GDL) a molecule  that slowly hydrlolyses into an acid and thus lowers slowly, continuously and homogeneously the pH of the medium, fig.2aXXX. This GDL controled acidification permits the formation of a bulk gel phase. We monitor the GDL induced gelation with a rheometer in cone plate geometry. As casein strongly adsorbs to surfaces, the strain and stress are well transmitted to the casein dispersion and we can measure the elatic and viscous modulus of the casein dispersion as a function of time. the mechanical properties of the gel are closely related to the fraction of free casein. First, the dispersion solidifies then reaches its maximum elasticity (500 Pa) at the isoelectric point and finaly the elastic modulus decreases down to 100 Pa.

To gain further insight into the mechanical properties of the gel we turn to micro structural observations under confocal microscopy. As shown in supplementary fig. 1XXX the gel consist in a space spaning fractal network with a curtoff length scale xi. xi corresponds to the maximum pore size of the network. This model fits the structure at all times but its parameters vary non monotoneously during GDL acidification. As shown in Fig. 2cXXX, up to the isolectric point xi increases up to a maximum length of 7 $\mu$m due to the solvent expulsion out of the casein network [refXXX]. This stiffen the arm of the network and is responsible for the increase of elasticity. Beyond the isoelectric point xi decreases down to 4$\mu$m. Casein regain solubility [refXXX], the strand network swell, the pore shrinks and the gel weakens.

As seen on the confocal measurements, the gel is porous and is permeable to the solvant.  This protertie is determinant as it allow the gel to absord and expel the solvant and therefore change its volume when sticky boundary are turned off. To quantify this effect, we have embeded the casein gel inside a capillary and deeped verticaly the capilary in an aqueous buffer [refXXX]. In the supplementary figure XXX, we observe that the aqueous buffer goes through the gel and rise in the tube due to capillary forces. Exploitation of this experiment give the permeability of the gel: 67000 nm$^2$.

Boundary conditions is the final parameter that is essential to form wrinkled patterns. The evolution of xi when sticky boundaries are on reflects the inner constrains of the gel:  xi increases, the stand of the nework become thinner and the gel wants to retract; then xi decreases, the stands of network swell and the gel wants to expend. When the casein adhesion to the upper boundary is turned off, the gel still not wrinkles because it remains attached to the bottom wall, but it detach from the top wall. In this situation we can monitor xi, fig. XXX. xi remains mostly constant and the major effect is repported on the volume occupied by the gel as will be repported below.

@@

Figure~\ref{fig:sideview}a shows confocal measurements in the case where both top and bottom walls have been coated, a configuration favorable to wrinkling. Up to 10 min the dispersion is homogeneous and fills the entire slit. The black margins correspond to the glass walls. Around 23 min, the gel forms and simultaneously becomes thiner and more concentrated - the fluorescence signal which is proportional to the casein concentration is brighter compared to the situation at 10min. This effect is the fingerprint of syn\ae{}resis: as gelation proceeds, the solvent is expeled on both the upper and lower sides. We describe sample as a porous elastic film of thickness $h$ sandwiched between two viscous layers of respective thicknesses $H_1$ and $H_2$ in a cell of thickness $e=H_1+h+H_2$ (Fig.~\ref{fig:sideview}b). Extracting the top and bottom surface of the gel phase allows for quantitative measurements of the relative volume (Fig.~\ref{fig:sideview}c) and excess surface area of the gel film (Fig.~\ref{fig:sideview}d) (see Supplementary Methods). As expected from the system design, the evolution of the volume is non-monotoneous. After reaching a minimum around 35\% of its initial value at 23 min, the volume increases. Shortly after,  at 36 min, the planar film destabilizes and starts to wrinkle in the $z$-direction creating excess area. This confirms that the instability is due to swelling. Around 38 min, the amplitude of the wrinkles reach the thickness of the slit. The size of the primary pattern is now settled Supplementary Fig. \ref{fig:patterns, fig:growth}. However swelling continues, the gel flattens on both top and bottom walls (44 min) and finally buckles back (53 min) to form the secondary pattern (1h21). When the excess volume saturates, the dynamics stops. The amount of GDL controls the ability of the gel to swell: at low GDL concentrations we only observe the primary pattern, at high GDL concentrations we have observe the formation of higher order patterns.

@@

\begin{figure}
	\includegraphics{sideview}
	\caption{Measurements and modeling. (a) Confocal $(x,z)$ cuts showing syn\ae{}resis, swelling, wrinkling and cascade buckling. The scale bar is \SI{100}{\micro\metre} (real size ratio). (b) Schematic side view of the cell. The constant thickness gel film (orange) is surrounded by water. Yellow lines show the path of transmitted light through the gel phase. A region of interest is highlighted in gray. (c-d) Confocal microscopy measurement of the volume of the gel phase relative to cell volume and excess area. Crosses correspond to times in (a).}
	\label{fig:sideview}
\end{figure}

@@

The determination of the characteristic length $\lambda$ of a wrinkling pattern is an old problem. Large deformations of thin sheets is governed by F\"{o}ppl-von Karman equations  refLaudauXXX,  essentially impossible to solve in analytical form. However using scaling and asymptotic arguments one can obtain the dominant wavelength of the wrinking patern and isolate the physically meaningful ingredients. An elastic film buckles due to excess area with respect to its boundaries. Wrinkling is a buckling hindered by a substrate: the mismatch of elastic properties between the film and the substrate selects the preferred wrinkling wavelength~\cite{Gough1940, Bijlaard1946}. $\lambda$ is then set by the competition between the flexural modulus of the film $B=E/(1-\nu^2)$ ($E$ Young modulus, $\nu$ Poisson ratio) and a transverse load $\sigma_\perp$. For small deflections of  amplitude $A$, $\lambda$ scales as~\cite{Vella2009,Kolinski2009}: 
%
\begin{equation}
\lambda \sim \left( A h^3 \frac{B}{\sigma_{\perp}} \right)^{1/4}.
\label{eq:lstar}
\end{equation}

@@

In our situation we now have to determine $\sigma_\perp$. We identify three possible candidates for the transverse load $\sigma_\perp$: the weight or buoyancy of the gel, the porous flow through the gel or the viscous flows in the surrounding water layers. We confirmed that gravity plays no significant role by successfully repeating the wrinkling patern in a cell held vertically, i.e. with the weight acting longitudinally rather than transversely.

@@

We consider two limit scenario that leads to wrinkling parterns. In a first limit, the film of gel sits without sticking to the bottom wall of the cell ($H_1\rightarrow0$). Creating a 2D blister requires flowing solvent of viscosity $\eta$ through the gel of permeability $\alpha$. Darcy law~\cite{Darcy1856} relates the pressure gradient, estimated as $\sigma_{\perp}/h$, to the volume flux per unit area $v = \frac{\alpha}{\eta}\frac{\sigma_{\perp}}{h} = \frac{dA}{dt}$. Injecting this expression in Eq.~(\ref{eq:lstar}) yields the wavelength of the Darcy mode as
\begin{equation}
\lambda_D \sim h^{1/2} \alpha^{1/4} \Upsilon^{1/4}
\end{equation}
where $\Upsilon = B\tau/\eta$ is a dimensionless viscoelastic factor quantifying the relative stiffness of the film and the solvent at a characteristic time $\tau$. This expression emphasizes the kinetic nature of the wavelength selection through Darcy flow. However we note that the wavelength of a generation of the pattern does not evolve with time (see Fig.~\ref{fig:dynamics}a and Sup. Fig.~\ref{fig:growth}) once the gel is touching both walls. This arrest may be due to bending rigidity~\cite{LeGoff2014}. Here the relevant time scale $\tau$ corresponds to the times it takes for the wrinkle amplitude to saturate (see Supplementary Method for the precise definition).

@@

In conclusion we have taken advantage of the properties of the protein casein and carefully ingineered the route to form permeable bio-gels that, in slit geometry,  produce original cascade of wrinkling patterns  when casein adhesion to the slit walls is turned off.  We developed a model that captures the dynamical origin of the constrains exerted on the gel and predict the wrinkling wavelength: Poiseuille flows of the solvent above and below the gel layer for long wavelengths and a new mechanism, Darcy flows of the solvent through the gel layer for short wavelengths. Such a model experiment, because it relies on pH induced charge stabilisation/destablisation, should be generalisable to different sort of proteins. As such it illustrate daily life problems such as wrinkles encountered while putting up wall paper and could set a benchmark to explore potential applications in food sience refXXX or in optical devices such as diffraction gratins or fresnel lenses refXXX, and the possibility of wrinkling in confined porous soft materials immersed in a buoyancy-matched viscous medium such as biological tissues refXXX.

@@

Sodium caseinate (Firmenich)  is disolved in deionised water by stiring for 20 min  at room temperature.  The mixing of the caseine dispersion with glucono-$\delta$-lactone (GDL) in powder (Firmenich) marks the begining of the experiments. We tipically have about 10 min to pipette the dispersion in the slit before gelation starts. Sodium caseinate (Firmenich) is labelled with Dylight~550 NHS ESTER (Thermo Scientific). Excess dye is removed by centrifugation. For confocal microscopy measurements, we add 10\% of fluorescently labelled caseines to the caseine/GDL dispersion while it is fluid.

Home made optical cells are composed of a slide (RS France) and a cover slip (Menzel-Gl\"aser) spaced by parafilm (Parafilm). The optical cell is heated  on a hot plate so that the parafilm welds to the  slide and the cover slip. Gently applying a pressure on the cover slip while being heated allows us to tune the thickness of the cell between 50 et 150 $\mu$m. The slit is filled with the casein dispersion, immediately sealed using ultraviolet-cured glue (Norland Optical) and placed under the microscope before gelation could take place. To prevent adsorption of caseins to surfaces, microscope slides and cover slips used to make microscopy cells are cleaned and subsequently coated with a silane agent (3-(trimethoxysilyl) propyl methacrylate, Sigma Aldrich) to which a polyacrylamide brush is polymerized from methacrylate groups (euromedex).

3D data are collected on a Zeiss LSM510 confocal microscope, using \SI{532}{\nano\meter} laser excitation. We use either a 10x (air) lens to measure the geometry of wrinkling (see Supplementary Material) or a 63x (oil) lens to observe the microstructure of the gel. Larger scope pictures, obtained either by stitching fluorescent microscopy images (Nikon Eclipse Ti) or by transmitted/reflected light macroscope (Nikon SMZ745T/Leica DMS1000) are used to measure the wavelength of the primary pattern (Supplementary Figures~\ref{fig:patterns} and \ref{fig:generations}). Secondary pattern wavelength is set to half of the primary wavelength, according to Supplementary Fig.~\ref{fig:generations}.
Data shown in Fig.~\ref{fig:acidification} a-b are obtained with a SevenCompact pH-meter (Mettler Toledo) and an MCR 301 rheometer (Anton Paar).
Permeability is measured using a protocol found in \cite{VanDijk1986} (see Supplementary Method and Supplementary Fig.~\ref{fig:permeability}). The amount of free casein protein present in the gel during the GDL acidificaition process was measured by titration and UV-spectrometry (Ocean Optics) according to the protocol found in [refXXX]

@@

the fraction of free casein proteins of a 100mL, 4\%w sodium caseinate dispersion is obtained either by titration by HCl 1 molar or during the action of 4\%w GDL. In both cases, 0.4mL of the 100mL dispersion is sampled regularly as a function of the pH, and centrifuged in eppendorf tubes at 14500 rpm for 3 min. The supernatent is diluted and analysed with a UV absorption spectrometer(Ocean Optics). According to refXXX, we deduced the fraction of free casein proteins from the adsorption spectrum:
%
\begin{equation}
x_{free}=\frac{A(\text{pH})}{A(\text{pH}=7)}\text{, with } A(\text{pH})=A_{280}-A_{320}XXX
\end{equation}
%
where  $A_{280}$ is the absortion peak due to XXX and $A_{320}$ is an estimate of the absorption due to turbidity.

@@

We first denoise each plane by a short-range Gaussian blur ($\sigma=\SI{2}{px}\ll\lambda$). Then, for each $(x,y)$, we focus on the z-dependent intensity profile $I(z)$. We define the position of the top surface of the gel $z_\text{top}$ as the point where $I(z)$ first crosses half of its maximum value over the given vertical line. Since this criterion is local, we do not need to correct for background intensity variation. In this way we obtain $z_\text{top}(x,y)$ with subpixel resolution ($\pm\SI{0.1}{px}=\SI{0.67}{\micro\metre}$), and $z_\text{bottom}(x,y)$ in a symmetrical way.

>>>>

\begin{abstract}
The slow acidification of caseins, a family of milk proteins, is well known to result in the formation of gels -- this is indeed one of the key steps to make yoghurt. Here, using the casein proteins as  building blocks and the acidulent Glucono-$\delta$-lactone, we design model thin biogel layers that wrinkles in cascade. Using a combination of titration, rheology, light microscopy and confocal microscopy, we demonstrate that, during the acidification, the gel shrinks and swells creating constrains that induce Darcy and Poiseuille flows and wrinkles the gel. We have develop an elasto-hydrodynamic model that predicts correctly the wrinkling wavelength. Our results should be universal for acid-induced protein gels as it is based on pH induced charge stabilisation/destabilisation and on control of boundary conditions. As such it could set a benchmark to gain fundamental insights into wrinkled biological tissues, to texture food or to design surfaces for optical purposes.
\end{abstract}

@@

There are many ways and reasons for a film to wrinkle. An elastic film buckles due to excess area with respect to its boundaries and wrinkles when buckling is hindered by a substrate. The selection of the preferred wrinkling wavelength is peculiar to the situation and is related to the mismatch of elastic properties between the surface and the substrate~\cite{Gough1940, Bijlaard1946, Biot1957, Bowden1998, Cerda2003, Genzer2006}.  On the biology side wrinkling-controlled morphogenesis is ubiquitous. Ageing and the loss of elastic fibres makes our skin wrinkles~\cite{Bissett1987, Genzer2006}. Difference in growth rates between the gut tube and its dorsal anchoring is responsible for the vilification of guts~\cite{Savin2011, Ciarletta2014, Shyer2013}. Localized cell death in biofilms focuses mechanical forces and initiates 3D labyrinth pattern~\cite{Trejo2013, Asally2012}. On the physics side the last two decades have seen a bloom of methods to obtain well-controlled patterns via linear~\cite{Bowden1998, Genzer2006, Hu1998, Kim2010, Vandeparre2011, Li2013} or nonlinear behaviour~\cite{Efimenko2005, Guvendiren2010, Kim2011a, Brau2011}. Such patterns can be triggered by temperature dilation~\cite{Bowden1998}, swelling~\cite{Hu1998} or the removal of pre-strain~\cite{Genzer2006}.

@@

Here, we introduce model permeable biogels that produce hierarchical wrinkling as a result of their interplay between their gelation dynamics and the confinement conditions. In slit geometry, we slowly acidify a dispersion of caseins, a family of milk proteins. As a result, not only does the casein form a gel as expected~\cite{Bremer1989} -- this is a well-known effect used in yoghurt making -- but the gel starts to produce wrinkling patterns in cascade when casein adhesion to the slit walls is turned off. Using a combination of rheology, light microscopy and confocal microscopy, we demonstrate that upon acidification the casein network spontaneously shrinks then swells at the micron level and wrinkles at the millimetre scale. We systematically investigate such swelling-induced wrinkling of a porous gel film sandwiched between two viscous layers, the whole system being confined between two glass plates. As shown in Fig.~\ref{fig:dynamics} and in Supplementary Movies 1 and 2, transverse confinement leads to saturation of the amplitude on both the top and bottom walls followed by a cascade of secondary and ternary buckling. Such cascade buckling has already been observed by~\cite{Roman1999} in a one-dimensional macroscopic system, a sheet of steel forced to buckle between two solid plates. In addition one has to note that the primary wavelength is much smaller than the size of the system, implying that a simple buckling scenario cannot explain the observations in Fig.~\ref{fig:dynamics} and that only a wrinkling instability can be invoked.

@@

\begin{figure}
	\includegraphics{acidification}%
	\caption{Acid-induced protein gel properties behave non monotonically with pH. (a) pH decrease over time in a 4\%w sodium caseinate solution acidified by 4\%w GDL. The horizontal line indicates isoelectric pH of caseins. (b) Evolution of the storage modulus $G'$ measured in a rheometer with full adhesion to rotor and stator. (c) Evolution of pore size measured by confocal microscopy in a slit geometry with either adhesion to all walls (black line) or no adhesion to top wall, i.e. allowed syn\ae{}resis and swelling (orange dashed line).}%
	\label{fig:acidification}%
\end{figure}

To form wrinkled patterns such as the one oberved in figure 1XXX, ones needs to take advantage of the pH sensitivity of caseins to control the route to gelation, to fine tune the gel properties and the boundary conditions.

The stability of caseins dispersion strongly depend on the pH of the solution. Supplementary movie 3XXX shows the titration of a caseins dispersion at 1\%w by 1 molar HCl. At pH=7, the caseins are well dispersed then then flocculate between pH= XXX to XXX and finaly at pH=XXX the solution become clear again. However as one increases the caseins concentration up to 4\%w, at low pH the caseins redispersion is not total, Supplementary movie 4XXX. The state diagram of caseins as a function of the pH  is related to charge stabilization. At neutral pH caseins are charged negatively and repel each other. As the pH decreases to the isoelectric point of the caseins (pH=4.5), caseins molecules loose their negative net charge and tend to stick to each other [ref XXX casein stability]. Beyond the isoelectric point casein acquire a positive net charge and therefore partially re-dissolve. In figure 2cXXX, we quantify this effect and measure with UV absorption spectroscopy the amount of free caseins within the 4\%w dispersion as a function of pH (see Supplementary Methods). Starting at pH=6.8 all the casein molecules are well dispersed and the free casein fraction is 1. Upon decreasing the pH of the solution to the isolectric point the free casein fraction drop to almost 0, almost all the casein molecule are aggregated. Further decrease of the pH of the solution lead to a partial redissolution of the casein aggregates. At pH=2 the fraction of free casein is XXX.

@@

To form a space spanning gel, flocculation need to be prevented. We use of Glucono-$\delta$-lactone (GDL) a molecule  that slowly hydrolyses into an acid and thus lowers slowly, continuously and homogeneously the pH of the medium, fig.2aXXX. This GDL controlled acidification permits the formation of a bulk gel phase. We monitor the GDL induced gelation with a rheometer in cone plate geometry. As casein strongly adsorbs to surfaces, the strain and stress are well transmitted to the casein dispersion and we can measure the elastic and viscous modulus of the casein dispersion as a function of time. the mechanical properties of the gel are closely related to the fraction of free casein. First, the dispersion solidifies then reaches its maximum elasticity (500 Pa) at the isoelectric point and finally the elastic modulus decreases down to 100 Pa.

To gain further insight into the mechanical properties of the gel we turn to micro structural observations under confocal microscopy. As shown in supplementary fig. 1XXX the gel consist in a space spanning fractal network with a cut off length scale xi. xi corresponds to the maximum pore size of the network. This model fits the structure at all times but its parameters vary non monotonously during GDL acidification. As shown in Fig. 2cXXX, up to the isolectric point xi increases up to a maximum length of 7 $\mu$m due to the solvent expulsion out of the casein network [refXXX]. This stiffen the arm of the network and is responsible for the increase of elasticity. Beyond the isoelectric point xi decreases down to 4$\mu$m. Casein regain solubility [refXXX], the strand network swell, the pore shrinks and the gel weakens.

As seen on the confocal measurements, the gel is porous and is permeable to the solvent.  This property is determinant as it allow the gel to absorb and expel the solvent and therefore change its volume when sticky boundary are turned off. To quantify this effect, we have embedded the casein gel inside a capillary and dip vertically the capillary in an aqueous buffer [refXXX]. In the supplementary figure XXX, we observe that the aqueous buffer goes through the gel and rise in the tube due to capillary forces. Exploitation of this experiment give the permeability of the gel: 67000 nm$^2$.

Boundary conditions is the final parameter that is essential to form wrinkled patterns. The evolution of xi when sticky boundaries are on reflects the inner constrains of the gel:  xi increases, the stand of the network become thinner and the gel wants to retract; then xi decreases, the stands of network swell and the gel wants to expend. When the casein adhesion to the upper boundary is turned off, the gel still not wrinkles because it remains attached to the bottom wall, but it detach from the top wall. In this situation we can monitor xi, fig. XXX. xi remains mostly constant and the major effect is reported on the volume occupied by the gel as will be reported below.

@@

Figure~\ref{fig:sideview}a shows confocal measurements in the case where both top and bottom walls have been coated, a configuration favourable to wrinkling. Up to 10 min the dispersion is homogeneous and fills the entire slit. The black margins correspond to the glass walls. Around 23 min, the gel forms and simultaneously becomes thinner and more concentrated - the fluorescence signal which is proportional to the casein concentration is brighter compared to the situation at 10min. This effect is the fingerprint of syn\ae{}resis: as gelation proceeds, the solvent is expelled on both the upper and lower sides. We describe sample as a porous elastic film of thickness $h$ sandwiched between two viscous layers of respective thickness $H_1$ and $H_2$ in a cell of thickness $e=H_1+h+H_2$ (Fig.~\ref{fig:sideview}b). Extracting the top and bottom surface of the gel phase allows for quantitative measurements of the relative volume (Fig.~\ref{fig:sideview}c) and excess surface area of the gel film (Fig.~\ref{fig:sideview}d) (see Supplementary Methods). As expected from the system design, the evolution of the volume is non-monotonous. After reaching a minimum around 35\% of its initial value at 23 min, the volume increases. Shortly after,  at 36 min, the planar film destabilizes and starts to wrinkle in the $z$-direction creating excess area. This confirms that the instability is due to swelling. Around 38 min, the amplitude of the wrinkles reach the thickness of the slit. The size of the primary pattern is now settled Supplementary Fig. \ref{fig:patterns, fig:growth}. However swelling continues, the gel flattens on both top and bottom walls (44 min) and finally buckles back (53 min) to form the secondary pattern (1h21). When the excess volume saturates, the dynamics stops. The amount of GDL controls the ability of the gel to swell: at low GDL concentrations we only observe the primary pattern, at high GDL concentrations we have observe the formation of higher order patterns.

@@

\begin{figure}
	\includegraphics{sideview}
	\caption{Measurements and modelling. (a) Confocal $(x,z)$ cuts showing syn\ae{}resis, swelling, wrinkling and cascade buckling. The scale bar is \SI{100}{\micro\metre} (real size ratio). (b) Schematic side view of the cell. The constant thickness gel film (orange) is surrounded by water. Yellow lines show the path of transmitted light through the gel phase. A region of interest is highlighted in gray. (c-d) Confocal microscopy measurement of the volume of the gel phase relative to cell volume and excess area. Crosses correspond to times in (a).}
	\label{fig:sideview}
\end{figure}

@@

The determination of the characteristic length $\lambda$ of a wrinkling pattern is an old problem. Large deformations of thin sheets is governed by F\"{o}ppl-von Karman equations  refLaudauXXX,  essentially impossible to solve in analytical form. However using scaling and asymptotic arguments one can obtain the dominant wavelength of the wrinkling pattern and isolate the physically meaningful ingredients. An elastic film buckles due to excess area with respect to its boundaries. Wrinkling is a buckling hindered by a substrate: the mismatch of elastic properties between the film and the substrate selects the preferred wrinkling wavelength~\cite{Gough1940, Bijlaard1946}. $\lambda$ is then set by the competition between the flexural modulus of the film $B=E/(1-\nu^2)$ ($E$ Young modulus, $\nu$ Poisson ratio) and a transverse load $\sigma_\perp$. For small deflections of  amplitude $A$, $\lambda$ scales as~\cite{Vella2009,Kolinski2009}: 
%
\begin{equation}
\lambda \sim \left( A h^3 \frac{B}{\sigma_{\perp}} \right)^{1/4}.
\label{eq:lstar}
\end{equation}

@@

In our situation we now have to determine $\sigma_\perp$. We identify three possible candidates for the transverse load $\sigma_\perp$: the weight or buoyancy of the gel, the porous flow through the gel or the viscous flows in the surrounding water layers. We confirmed that gravity plays no significant role by successfully repeating the wrinkling pattern in a cell held vertically, i.e. with the weight acting longitudinally rather than transversely.

@@

We consider two limit scenario that leads to wrinkling patterns. In a first limit, the film of gel sits without sticking to the bottom wall of the cell ($H_1\rightarrow0$). Creating a 2D blister requires flowing solvent of viscosity $\eta$ through the gel of permeability $\alpha$. Darcy law~\cite{Darcy1856} relates the pressure gradient, estimated as $\sigma_{\perp}/h$, to the volume flux per unit area $v = \frac{\alpha}{\eta}\frac{\sigma_{\perp}}{h} = \frac{dA}{dt}$. Injecting this expression in Eq.~(\ref{eq:lstar}) yields the wavelength of the Darcy mode as
\begin{equation}
\lambda_D \sim h^{1/2} \alpha^{1/4} \Upsilon^{1/4}
\end{equation}
where $\Upsilon = B\tau/\eta$ is a dimensionless viscoelastic factor quantifying the relative stiffness of the film and the solvent at a characteristic time $\tau$. This expression emphasizes the kinetic nature of the wavelength selection through Darcy flow. However we note that the wavelength of a generation of the pattern does not evolve with time (see Fig.~\ref{fig:dynamics}a and Sup. Fig.~\ref{fig:growth}) once the gel is touching both walls. This arrest may be due to bending rigidity~\cite{LeGoff2014}. Here the relevant time scale $\tau$ corresponds to the times it takes for the wrinkle amplitude to saturate (see Supplementary Method for the precise definition).

@@

In conclusion we have taken advantage of the properties of the protein casein and carefully ingineered the route to form permeable bio-gels that, in slit geometry,  produce original cascade of wrinkling patterns  when casein adhesion to the slit walls is turned off.  We developed a model that captures the dynamical origin of the constrains exerted on the gel and predict the wrinkling wavelength: Poiseuille flows of the solvent above and below the gel layer for long wavelengths and a new mechanism, Darcy flows of the solvent through the gel layer for short wavelengths. Such a model experiment, because it relies on pH induced charge stabilisation/destabilisation, should be generalisable to different sort of proteins. As such it illustrate daily life problems such as wrinkles encountered while putting up wall paper and could set a benchmark to explore potential applications in food science refXXX or in optical devices such as diffraction gratings or Fresnel lenses refXXX, and the possibility of wrinkling in confined porous soft materials immersed in a buoyancy-matched viscous medium such as biological tissues refXXX.

@@

Sodium caseinate (Firmenich)  is dissolved in deionised water by stirring for 20 min  at room temperature.  The mixing of the caseine dispersion with glucono-$\delta$-lactone (GDL) in powder (Firmenich) marks the beginning of the experiments. We typically have about 10 min to pipette the dispersion in the slit before gelation starts. Sodium caseinate (Firmenich) is labelled with Dylight~550 NHS ESTER (Thermo Scientific). Excess dye is removed by centrifugation. For confocal microscopy measurements, we add 10\% of fluorescently labelled caseins to the casein/GDL dispersion while it is fluid.

Home made optical cells are composed of a slide (RS France) and a cover slip (Menzel-Gl\"aser) spaced by parafilm (Parafilm). The optical cell is heated  on a hot plate so that the parafilm welds to the  slide and the cover slip. Gently applying a pressure on the cover slip while being heated allows us to tune the thickness of the cell between 50 et 150 $\mu$m. The slit is filled with the casein dispersion, immediately sealed using ultraviolet-cured glue (Norland Optical) and placed under the microscope before gelation could take place. To prevent adsorption of caseins to surfaces, microscope slides and cover slips used to make microscopy cells are cleaned and subsequently coated with a silane agent (3-(trimethoxysilyl) propyl methacrylate, Sigma Aldrich) to which a polyacrylamide brush is polymerized from methacrylate groups (Euromedex).

3D data are collected on a Zeiss LSM510 confocal microscope, using \SI{532}{\nano\meter} laser excitation. We use either a 10x (air) lens to measure the geometry of wrinkling (see Supplementary Material) or a 63x (oil) lens to observe the microstructure of the gel. Larger scope pictures, obtained either by stitching fluorescent microscopy images (Nikon Eclipse Ti) or by transmitted/reflected light macroscope (Nikon SMZ745T/Leica DMS1000) are used to measure the wavelength of the primary pattern (Supplementary Figures~\ref{fig:patterns} and \ref{fig:generations}). Secondary pattern wavelength is set to half of the primary wavelength, according to Supplementary Fig.~\ref{fig:generations}.
Data shown in Fig.~\ref{fig:acidification} a-b are obtained with a SevenCompact pH-meter (Mettler Toledo) and an MCR 301 rheometer (Anton Paar).
Permeability is measured using a protocol found in \cite{VanDijk1986} (see Supplementary Method and Supplementary Fig.~\ref{fig:permeability}). The amount of free casein protein present in the gel during the GDL acidification process was measured by titration and UV-spectrometry (Ocean Optics) according to the protocol found in [refXXX]

@@

the fraction of free casein proteins of a 100mL, 4\%w sodium caseinate dispersion is obtained either by titration by HCl 1 molar or during the action of 4\%w GDL. In both cases, 0.4mL of the 100mL dispersion is sampled regularly as a function of the pH, and centrifuged in eppendorf tubes at 14500 rpm for 3 min. The supernatant is diluted and analysed with a UV absorption spectrometer(Ocean Optics). According to refXXX, we deduced the fraction of free casein proteins from the adsorption spectrum:
%
\begin{equation}
x_{free}=\frac{A(\text{pH})}{A(\text{pH}=7)}\text{, with } A(\text{pH})=A_{280}-A_{320}XXX
\end{equation}
%
where  $A_{280}$ is the absortion peak due to XXX and $A_{320}$ is an estimate of the absorption due to turbidity.

@@

We first de-noise each plane by a short-range Gaussian blur ($\sigma=\SI{2}{px}\ll\lambda$). Then, for each $(x,y)$, we focus on the z-dependent intensity profile $I(z)$. We define the position of the top surface of the gel $z_\text{top}$ as the point where $I(z)$ first crosses half of its maximum value over the given vertical line. Since this criterion is local, we do not need to correct for background intensity variation. In this way we obtain $z_\text{top}(x,y)$ with subpixel resolution ($\pm\SI{0.1}{px}=\SI{0.67}{\micro\metre}$), and $z_\text{bottom}(x,y)$ in a symmetrical way.

From these data, we obtain for each stack the volume and thickness of the gel phase, area of the mid surface and peak-to-peak amplitude. Velocity is obtained by time differentiation. Once the instability has saturated on the top and bottom walls of the cell, peaks in the histograms of $z_\text{top}$ and $z_\text{bottom}$ yield the measure of the positions of both walls, thus of the cell thickness $e$ and retrospectively of $H_1$ and $H_2$.
